---
title: "Project Description"
author: Rhea Bhat, Yaashi Khatri, Ian Wood, Neha George, Natalie Nguyen, Andrea Flores,
  Hiba Ansari
date: "2022-09-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Description: Estimation of Obesity Levels

#### Rhea Bhat, Yaashi Khatri, Ian Wood, Neha George, Natalie Nguyen, Andrea Flores, Hiba Ansari
#### rb43853, ha9982, nmn498, nmg2337 ADD YOUR EIDS HERE

### Project Goals
We want to examine the effect of eating habit and physical conditions on obesity levels in Mexico, Peru, and Colombia. We will utilize a multiple regression model to examine these effects. 

### Background

**Factors Contributing to Obesity**

Obesity is a disease characterized by an excessive accumulation of body fat that poses a risk to the individual's health. The tool for measuring whether an individual is overweight or obese is through body mass index (BMI), in which a BMI of 25-29.9 kg/m^2^ indicates being overweight and a BMI  $>$ 30 indicates obesity (Fock & Khoo, 2013). There are a multitude of factors that contribute to the development of obesity. Among these factors include diet, physical conditions, and family history. 

- __*Diet*__
   
    An imbalanced diet is the main contributing factor to the onset of obesity. Individuals that subscribe to diets with higher intakes in saturated and trans fat, refined grains, red meats, sweets, and sweetened sugary-beverages are observed to have increased weight gain and risk of obesity. Conversely, individuals with diets high in whole grains, vegetables, fruit, yogurt, nuts, fish, and poultry were observed to have less weight gain and decreased risk of obesity (Hruby et al., 2016).


- __*Physical Conditions*__
    
    Physical conditions refer to exercise, sedentary vs. active lifestyle, and sleep quality. According to a study conducted by Hruby et al. (2016), higher levels of physical activity (defined as activity $\ge$ 30 min/day) resulted in less weight gained than low physical activity levels (defined as activity $\le$ 30 min/day). Conversely, greater times spent engaged in sedentary activities (watching TV, using cellphone, playing video games, etc.) results in increased risk of obesity. The built environment of individuals may influence BMI as well. Higher density areas that are more walkable are likely to promote walking or bicycling as a means of transportation, which would increase average physical activity in these areas (Hruby et al., 2016).
    
    Sleep duration and quality also impacts risk of weight gain. It was found that short duration sleep (time slept $\le$ 6 hr per night) and poor sleep quality are linked to increased risk of obesity (Beccuti & Pannain, 2011).

    
- __*Family History*__
   
    Family history also contributes to the development of obesity and consists genetic factors and home environment. Genetics may impact dietary palate, food sensitivity, and metabolism of individuals. Variations in these genes could result in adverse eating choices (i.e., selective, limited, or excessive intake of different food groups). Moreover, Families of varying ethnic and cultural backgrounds are exposed to different types of diets that emphasize different food groups. Family environment also influences eating behaviors, which determines diet quality and quantity. The home environment determines if food is provided consistently, prepared at home, and if it is healthy and nutritious (Fruh, 2017).
    
    
    
#### References
Beccuti, G., & Pannain, S. (2011). Sleep and obesity: Current Opinion in Clinical Nutrition and Metabolic Care, 14(4), 402–412. https://doi.org/10.1097/MCO.0b013e3283479109

Fock, K. M., & Khoo, J. (2013). Diet and exercise in management of obesity and overweight: Diet and exercise for weight management. Journal of Gastroenterology and Hepatology, 28, 59–63. https://doi.org/10.1111/jgh.12407

Fruh, S. M. (2017). Obesity: Risk factors, complications, and strategies for sustainable long‐term weight management. Journal of the American Association of Nurse Practitioners, 29(S1), S3–S14. https://doi.org/10.1002/2327-6924.12510

Hruby, A., Manson, J. E., Qi, L., Malik, V. S., Rimm, E. B., Sun, Q., Willett, W. C., & Hu, F. B. (2016). Determinants and consequences of obesity. American Journal of Public Health, 106(9), 1656–1662. https://doi.org/10.2105/AJPH.2016.303326



### Task
The purpose of our project is to predict the obesity levels based on various eating habits and physical condition. In order to analyze our data, we will be using regression.

### Dataset
The set includes data for the estimation of obesity levels in individuals from the countries of Mexico, Peru and Colombia, based on their eating habits and physical condition. The data contains 17 attributes and 2111 records, the records are labeled with the class variable `NObesity` (Obesity Level), which allows classification of the data using the values of Insufficient Weight, Normal Weight, Overweight Level I, Overweight Level II, Obesity Type I, Obesity Type II and Obesity Type III. 

### Visualization

```{r message=FALSE}
#Loading the data
library(tidyverse)
library(readxl)
obesityData <- read_excel("~/proj/obesityData.xlsx")
View(obesityData)
```

#### Before creating a visualization, here is a glimpse of the data

```{R}
# Converting
odata <- as.data.frame(obesityData)
head(odata)
```

#### Below are a few visualizations that help us better understand the data

```{R}
odata %>% group_by(Age) %>% ggplot(aes(x=NObeyesdad, y=Weight, color=Gender)) + geom_bar(stat="identity") + xlab("Weight Classification") + ylab("Weight (in lbs) ")+ ggtitle("Bar Plot: Weight vs. Classification") + scale_x_discrete(guide = guide_axis(n.dodge=2))
```

